---
title: "The relative popularity of programming languages between 2008 and 2018" 
author: "HUYNH DO MINH PHUONG"
date: "2/28/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Data on tags over time

How can we tell what programming languages and technologies are used by the most people? How about what languages are growing and which are shrinking, so that we can tell which are most worth investing time in? 

One excellent source of data is Stack Overflow, a programming question and answer site with more than 16 million questions on programming topics. By measuring the number of questions about each technology, we can get an approximate sense of how many people are using it. We're going to use open data from the Stack Exchange Data Explorer to examine the relative popularity of languages like R, Python, Java and Javascript have changed over time. 

Each Stack Overflow question has a tag, which marks a question to describe its topic or technology. For instance, there's a tag for languages like R or Python, and for packages like ggplot2 or pandas. 

We'll be working with a dataset with one observation for each tag in each year. The dataset includes both the number of questions asked in that tag in that year, and the total number of questions asked in that year.

```{r}
# Load libraries
library(tidyverse)
```

```{r}
# Loading the dataset
by_tag_year <- read_csv("by_tag_year.csv")
```

```{r}
# Inspect the dataset
str(by_tag_year)
```

<p style="color:blue"><i> The data has 40.518 rows and 4 columns corresponding to the following information:</i></p>
<p style="color:blue"><i> - 'year' : 10 years from 2008 to 2018,</i></p>
<p style="color:blue"><i> - 'tag': types of programming languages in the past 10 years,</i></p>
<p style="color:blue"><i> - 'number' : the number of questions with tag during year,</i></p>
<p style="color:blue"><i> - 'year_toal' : total number of questions on all programming topics in the year.
</i></p>


## 2. Now in fraction format

This data has one observation for each pair of a tag and a year, showing the number of questions asked in that tag in that year and the total number of questions asked in that year. For instance, there were 54 questions asked about the .htaccess tag in 2008, out of a total of 58390 questions in that year. 

Rather than just the counts, we're probably interested in a percentage: the fraction of questions that year that have that tag. So let's add that to the table.

```{r}
library(lubridate) # Load lubridate to make dealing with dates
# Add fraction column and convert num to date
by_tag_year_fraction <- by_tag_year %>% 
  mutate(fraction = number/year_total)
  #mutate(year = make_date(year)) 
```

```{r}
# Print the new table
by_tag_year_fraction
```


## 3. Has R been growing or shrinking?

So far we've been learning and using the R programming language. Wouldn't we like to be sure it's a good investment for the future? Has it been keeping pace with other languages, or have people been switching out of it? 

Let's look at whether the fraction of Stack Overflow questions that are about R has been increasing or decreasing over time.

```{r}
# Filter for R tags
r_over_time <- by_tag_year_fraction %>% 
  filter(tag == "r")
```

```{r}
# Print the new table
r_over_time 
```


## 4. Visualizing change over time

Rather than looking at the results in a table, we often want to create a visualization. Change over time is usually visualized with a line plot.

```{r}
# Create a line plot of fraction over time
r_over_time %>% 
  ggplot(aes(year,fraction)) + geom_line(colour = "blue") +
  labs(title = "CHANGE OF 'R' TAG OVER TIME", x = "Year", y = "Fraction") +
      theme(panel.background = element_rect(fill = "GhostWhite") ) +
  scale_x_continuous(breaks = 0:2100)
```

<p style="color:blue"><i> Interest in the R programming language has been grown rapidly over the past 10 years of the ever-increasing number of questions on this topic.</i></p>


## 5. How about dplyr and ggplot2?

Based on that graph, it looks like R has been growing pretty fast in the last decade. Good thing we're practicing it now! 

Besides R, two other interesting tags are dplyr and ggplot2, which we've already used in this analysis. They both also have Stack Overflow tags! 

Instead of just looking at R, let's look at all three tags and their change over time. Are each of those tags increasing as a fraction of overall questions? Are any of them decreasing?

```{r}
# A vector of selected tags
selected_tags <- c("r", "dplyr","ggplot2")
```

```{r}
# Filter for those tags
selected_tags_over_time <- by_tag_year_fraction %>% 
    filter(tag %in% selected_tags)
```

```{r}
# Plot tags over time on a line plot using color to represent tag
selected_tags_over_time %>%
  ggplot(aes(year,fraction,color = tag)) + geom_line() +
  labs(
    title = "CHANGE OF R, DPLYR AND GGPLOT2 TAGS OVER TIME", x = "Year", y = "Fraction") +
    theme(panel.background = element_rect(fill = "GhostWhite") ) +
  scale_x_continuous(breaks = 0:2100)
```

<p style="color:blue"><i> R has grown quite rapidly over the last decade while later languages like dplyr and ggplot2 have a slight increase.</i></p>


## 6. What are the most asked-about tags?

It's sure been fun to visualize and compare tags over time. The dplyr and ggplot2 tags may not have as many questions as R, but we can tell they're both growing quickly as well.

We might like to know which tags have the most questions overall, not just within a particular year. Right now, we have several rows for every tag, but we'll be combining them into one. That means we want group_by() and summarize().

Let's look at tags that have the most questions in history.

```{r}
# Find total number of questions for each tag
sorted_tags <- by_tag_year %>%
  group_by(tag) %>%
  summarize(number_total = sum(number)) %>%
  arrange(desc(number_total))
```

```{r}
# Print the new table
sorted_tags 
```


## 7. How have large programming languages changed over time?

We've looked at selected tags like R, ggplot2, and dplyr, and seen that they're each growing. What tags might be shrinking? A good place to start is to plot the tags that we just saw that were the most-asked about of all time, including JavaScript, Java and C#.

```{r}
# Get the six largest tags
highest_tags <- head(sorted_tags$tag)
```

```{r}
# Filter for the six largest tags
by_highest_tag_subset <- by_tag_year_fraction %>% 
    filter(tag %in% highest_tags)
by_highest_tag_subset
```

```{r}
# Plot tags over time on a line plot using color to represent tag
by_highest_tag_subset %>% 
  ggplot(aes(year,fraction,color = tag)) + geom_line() +
  labs(
    title = "TOP 6 HIGHEST TAGS OVER TIME", x = "Year", y = "Fraction") +
    theme(panel.background = element_rect(fill = "GhostWhite") ) +
  scale_x_continuous(breaks = 0:2100)
```

<p style="color:blue"><i> In the top 6 highest tags, python is the only asked tag that has steadily increased over the past decade, whereas c# is in a serious decline about number of questions. android, java and php rose sharply in the first 6 years and gradually decrease in the last 4 years, while javascript tended to reduce in the last 2 years after a strong increase from 2008.</i></p>


## 8. Some more tags!

Wow, based on that graph we've seen a lot of changes in what programming languages are most asked about. C# gets fewer questions than it used to, and Python has grown quite impressively. 

This Stack Overflow data is incredibly versatile. We can analyze any programming language, web framework, or tool where we'd like to see their change over time. Combined with the reproducibility of R and its libraries, we have ourselves a powerful method of uncovering insights about technology. 

To demonstrate its versatility, let's check out how three big mobile operating systems (Android, iOS, and Windows Phone) have compared in popularity over time. But remember: this code can be modified simply by changing the tag names!

```{r}
# Get tags of interest
my_tags <- c("android", "ios", "windows-phone" )
```

```{r}
# Filter for those tags
by_tag_subset <- by_tag_year_fraction %>% 
  filter(tag %in% my_tags)
by_tag_subset
```

```{r}
# Plot tags over time on a line plot using color to represent tag
by_tag_subset %>% 
  ggplot(aes(year,fraction,color = tag)) + geom_line() +
  labs(
    title = "CHANGE OF ANDROID, IOS, WINDOWS-PHONE OVER TIME", x = "Year", y = "Fraction") +
    theme(panel.background = element_rect(fill = "GhostWhite") ) +
  scale_x_continuous(breaks = 0:2100)
```

<p style="color:blue"><i> In the three main mobile operating systems (Android, iOS and Windows Phone), the Windows Phone tag has fewest question, whereas Android is the most popular language by fraction of questions increasing pretty fast in the last decade.</i></p>


# BONUS

## 9. Has 'R' been growing or shrinking?

```{r}
# Filter for R tags
r_programming <- by_tag_year_fraction %>% 
  filter(str_detect(tag, "^r\\b")) %>%
  group_by(year) %>%
  summarize(fraction_total = sum(fraction))
```

```{r}
# Create a line plot of fraction over time
r_programming %>% 
  ggplot(aes(year,fraction_total)) + geom_line(colour = "blue") +
  labs(
    title = "CHANGE OF R PROGRAMMING LANGUAGE OVER TIME", x = "Year", y = "Fraction") +
    theme(panel.background = element_rect(fill = "GhostWhite") ) +
  scale_x_continuous(breaks = 0:2100)
```


## 10. How about three big mobile operating systems (Android, iOS, and Windows Phone)

```{r}
android <- by_tag_year_fraction %>% 
  filter(str_detect(tag, "^android\\b")) %>%
  group_by(year) %>%
  summarize(android = sum(fraction))

ios <- by_tag_year_fraction %>% 
  filter(str_detect(tag, "^ios\\b")) %>%
  group_by(year) %>%
  summarize(ios = sum(fraction))

winphone <- by_tag_year_fraction %>% 
  filter(str_detect(tag, "^windows-phone\\b")) %>%
  group_by(year) %>%
  summarize(winphone = sum(fraction))

android_ios <- merge(android, ios, by="year", sort =F, all.x= T)

mobile_os <- merge(android_ios, winphone, by="year", sort=F, all.x=T) %>%
  gather(tag, total_fraction, android:winphone)
```

```{r}
# Plot tags over time on a line plot using color to represent tag
mobile_os %>% 
  ggplot(aes(year,total_fraction,color = tag)) + geom_line() +
  labs(
    title = "CHANGE OF 3 BIG MOBILE OPERATING SYSTEMS OVER TIME", x = "Year", y = "Fraction") +
    theme(panel.background = element_rect(fill = "GhostWhite") ) +
  scale_x_continuous(breaks = 0:2100)
```

<p style="color:blue"><i> There are no questions regarding the Windows Phone tag in 2009.</i></p>


## 11. How have top programming languages changed since 2014?

```{r}
# Filter for the tags since 2014 with fraction over 0.05
by_tag_top <- by_tag_year_fraction %>% 
  filter(year >= 2014, fraction > 0.05) %>%
  arrange(desc(fraction))
```

```{r}
library(gcookbook) # Load gcookbook for the cabbage_exp data set
# Plot tags over time on a stacked bar graph using color to selected tags
  ggplot(by_tag_top, aes(year, fraction, fill = tag)) +
  #geom_col() +
  geom_col(position = "fill") +
  labs(title = "CHANGE OF HIGHEST TAGS SINCE 2014", x = "Year", y = "Fraction") +
    theme(panel.background = element_rect(fill = "GhostWhite") ) +
  scale_x_continuous(breaks = 0:2100)
```

<p style="color:blue"><i> The number of questions asked in 'html' and 'jquery' tags after the year 2016 was gone, it means that those languages are shrinking. In the contrast, javascript and python are the programming languages that increases rapidly.</i></p>


# 12. How has average fraction of tag changed since 2014?

```{r}
# Filter for the tags since 2014 and count the tag
filter_1 <- by_tag_year_fraction %>% 
  filter(year >= 2014) %>%
  group_by(tag) 

filter_2 <- by_tag_year_fraction %>% 
  filter(year >= 2014) %>%
  count(tag)

by_tag_year_filter <- merge(filter_1, filter_2, by="tag", sort=F, all.x=T)
```

```{r}
filter_top <- by_tag_year_filter %>%  
  group_by(tag) %>%
  summarize(avg = sum(fraction)/mean(n)) %>%
  filter(avg >= 0.05) %>%
  arrange(desc(avg))
```

```{r}
# Plot tags over time on a Cleveland dot plot to filtered tags
ggplot(filter_top, aes(x = avg, y = reorder(tag, avg), col = tag)) +
  geom_point(size = 3) +  # Use a larger dot
  theme_classic () +
  theme(panel.grid.major.y = element_line(colour = "grey60", linetype = "dashed")) +
  labs(title = "AVERAGE FRACTION BY TAGS SINCE 2014", x = "Average Fraction", y = "Tag") 
```

<p style="color:blue"><i> Since 2014, the popularity of languages like Javascript and Java are still the most, whereas the number of people using c# and html is shrinking.</i></p>
